\chapter{Introduction}

Multiplayer networks in gaming have been around since the early days of gaming, but they have come a long way since then. In the past, multiplayer games were typically limited to local area networks or split-screen play, which required players to be physically present in the same location. However, with the advent of the internet, multiplayer networks have become more accessible and sophisticated.
The popularity of online gaming has fueled the development of multiplayer networks capable of supporting enormous numbers of players from all over the world. As a result, various types of multiplayer networks have emerged, each with its own set of pros and limitations.

The purpose of this minor dissertation is to investigate the use of multiplayer networking systems in multiplayer games, with a higher focus on hybrid server connections. The applied project portion of this dissertation will specifically involve creating a simple multiplayer game that communicates via a hybrid network system which assesses its effectiveness in terms of gameplay experience, network performance, and cost-effectiveness.

I believe this project will throw light onto some of the inner workings of multiplayer systems in online gaming, as well as provide myself and hopefully the reader some insight into how small developers choose between network structures and how the industry as a whole may have to evolve to allow for emerging technologies to take the stage and enable players to experience games like never seen before.

This document will be laid out as follows:

\begin{itemize}
    \item Methodology: This chapter describes my approach to testing and project planning, as well as some research tools I employed to get an understanding of multiplayer game infrastructures.
    \item Technology Review: provides a thorough examination of the many network architecture topologies commonly utilised in popular games, as well as new technologies that may impact the future of gaming.
    \item System Design: Documents the overall design of my project and how each component interacts with the rest. This goes into detail about some of the technologies used to develop the application.
    \item System Evaluation: Displays the overall effectiveness of the developed system, where strengths and weaknesses have risen and how they may be overcome with further research
    \item Conclusion: Highlights the findings of the research conducted, experiences encountered through development and any insights that may help with further development.
\end{itemize}



